Spontaneous breaking of multipole symmetries

### 2021-12-13 multipoleSSB/manuscript_Dec10.tex
comments from Rahul

The multipole group is well-explained in Ref.~\cite{Gromov2019}. For concreteness, and since we will be interested in situations where a multipolar symmetry group is spontaneously broken, we will describe the multipole group in terms of its action on a compact scalar field $\phi$, which will play the role of a Goldstone boson for the broken multipolar symmetry. We will always imagine $\phi$ as constituting the phase mode for some microscopic order parameter $\psi \sim e^{i\phi}$, with $\psi$ transforming by various $U(1)$ phases under the multipole group. \ethan{are there any examples showing that this interpretation is too restrictive?}

 The multipole group generalizes the internal shift symmetry $\phi(x) \goesto\phi (x) +c$ by allowing a shift by some set of polynomials involving the spatial coordinates, viz. $\phi (x) \goesto \phi (x) + \lambda_\alpha P^\alpha(x)$. The variables $\lambda_\alpha$ are symmetry parameters, while $\alpha$ labels the set of polynomials $P^\alpha(x)$. These so-called polynomial shift symmetries~\cite{Griffin2015} all commute with each other. It is helpful to limit ourselves to homogeneous polynomials. We can label these as $P_a^{I_a}$, where $a$ is the degree of the polynomial and $I_a$ is an abstract index that runs over the polynomials of degree $a$.

@@

Reference~\cite{Gromov2019} includes discussion of some multipole groups; we will review a few here. The simplest case is the maximal multipole group $\mathcal{M}^a_\text{max}$, which includes all shifts by polynomials of degree $a$ or less. Individual monomials can be written as
\begin{align}
P_c^{\gamma} = \mu^{\gamma}_{I_c}x^{I_c}, \label{eqn:basis}
\end{align}
where $\gamma$ is an index that runs over monomials of degree $c$ and $c\le a$. The composite index $I_c = \{i_1,\dots,i_c\}$  expands as $x^{I_c}=x^{i_1}\cdots x^{i_c}$, and similarly for $\partial_{I_c}$. Each  matrix  $\mu^{\gamma}$ is fully symmetric on its $I_c$ indices. This group also includes all translations and rotations.

@@

We can now consider building IR field theories for phases in which a multipolar symmetry is completely spontaneously broken, with no residual unbroken subgroup. Reference~\cite{Gromov2019} describes the process in detail, so in the following we will be somewhat succinct. 
We will continue to write things down in terms of the compact scalar $\phi$, which (at in the cases for which SSB is not preempted by strong fluctuations) is to be viewed as the Goldstone for the sponatneously broken symmetry.

@@

Furthermore, the structure of the quantum correlation function,
\begin{align}
C &\sim \int \frac{d\omega d^dk}{\omega^2 + K_k}\nn
&\sim \int \frac{d^dk}{\sqrt{K_k}},
\end{align}
suggests that, at least in some broad class of theories, the quantum critical dimension will be half the classical critical dimension. This is not true in general, though, as we will show for theories with quenched disorder.

@@

Consider a situation in which a dipolar symmetry $\mcm^1_{\rm max}$ is spontaneously broken to its monopolar subgroup $\mcm^0$, and momentarily specialize to the case of $d=3$ and $T>0$. In the spirit of Goldstone's theorem, the most naive thing to do when analyzing the symmetry broken phase would be to write down an action involving a set of $d=3$ Goldstone fields $\theta_j$, with $\t_j$ transforming linearly under the dipolar part of the symmetry group, but remaining invariant under the unbroken monopole subgroup, so that $\theta_j \mapsto \theta_j + \beta_j$ under a transformation parameterized by $P(x) = \alpha + \beta_j x^j$. In this case, the IR field theory for the putative symmetry-breaking phase would be of the form 
\begin{align}
\label{vector} \mathcal{L}= \sum_j  (\partial_\tau \theta_j)^2 + \sum_{i,j} g_{ij} (\partial_i \theta_j)^2.
\end{align}
As we are working in $d=3$ and at $T>0$, the theory \eqref{vector} indeed has SSB for the dipole symmetry, with $e^{i\t_j}$ developing long-range order.

@@

Let $b\leq a$ denote the smallest degree of multipole charge with nonzero compressibility, i.e. consider a situation in which $\kappa_d>0$ for all $b\leq d\leq a$, while $\kappa_d=0$ for all $d<b$. Then the minimal IR Lagrangian for the symmetry broken phase looks schematically like 
\be \label{generall} \mathcal{L} = \kappa_b (\partial_\tau \phi_{I_b})^2 + g_{J_{a-b+1}}(\partial_{J_{a-b+1}}\phi_{I_b})^2,\ee 
where 
$\p_{J_{a-b+1}} \equiv \p_{j_1} \cdots \p_{j_{a-b+1}}$, and where $\phi_{I_b}$ is a field that shifts by a polynomial of degree $a-b$ under the action of $\mmax a$. The compressibility appears as the coefficient of the time derivative term since $\p_\tau \phi_{I_b}$ is the momentum conjugate to $\phi_{I_b}$ in the quantum theory, and hence correlation functions of fluctuations in $n_b$ are determined by those of $\p_\tau \phi_{I_b}$.

A necessary condition for $\mmax b$ to be spontaneously broken is that $\kappa_b \neq 0$. Indeed, if $\mmax b$ is spontaneously broken then $n_b$ creates Goldstones when acting on the ground state, thereby ensuring that the density-density correlator appearing in the calculation of \eqref{kappadef} must be nonzero at long wavelengths. \ethan{is this too laconic?}

@@

In the previous subsection, we saw that the number of Goldstone modes in the symmetry-breaking theory is determined by the smallest degree of multipole charge $b$ such that $\kappa_b\neq0$. We now turn to determining the symmetry breaking pattern that occurs for a given choice of $b,d,$ and $T$, thereby allowing us to formulate a generalized version of the Mermin-Wagner theorem for the case of partial multipole breaking.

Since $\kappa_d=0$ if $d<b$, none of the subgroups $\mmax {d<b}$ can be spontaneously broken. Suppose then that the symmetry is spontaneously broken down to $\mmax {c}$, where $c\geq b$. To find the allowed values of $c$ consistent with this assumption, we need to calculate correlation functions of operators of the form $\exp(i\p_{J_{c-b}}\phi_{I_b})$, these being the operators transforming nontrivially under $\mmax c$ with the longest-range correlation functions. We have 
\be\ba \langle \p_{J_{c-b}}\phi_{I_b}(x) \p_{J_{c-b}}\phi_{I_b}(0)\rangle & \sim \int \frac{d\omega \, dk\, k^{d-1+2(c-b)}}{\omega^2 + k^{2(a-b+1)}} \\ 
& \sim \int dk\, k^{d-1+2(c-b) - \varsigma_T(a-b+1)}, \ea\ee 
where we have defined the function $\varsigma_T$ such that $\varsigma_T = 1$ if $T=0$ and $\varsigma_T = 2$ if $T>0$. Consequently, we find that $\mmax c$ can be spontaneously broken only in dimensions $d$ such that 
\be \label{pbreak_mw} d>2(b-c) + \varsigma_T(a-b+1).\ee 
For a given choice of symmetry breaking pattern (captured by $a$ and $c$) and a given choice of which compressibilities are nonzero (captured by $b$), the quantity on the RHS determines the lower critical dimension for symmetry breaking, thereby generalizing the Mermin-Wagner theorem to the partial breaking case. 
As a sanity check on \eqref{pbreak_mw}, note that when we set $c=b=0$ (full multipole breaking) we recover the lower critical dimension of $\varsigma_T(a+1)$ obtained in the previous section. Also note that as expected, the lower critical dimensions in the quantum $(\varsigma_T = 1)$ and classical $(\varsigma_T = 2)$ cases are related by $d_{\rm cl} = d_{\rm q} + z$, where $z=a-b+1$ is the dynamical exponent of the Goldstone theory \eqref{generall}.

@@

In a disordered classical system, the energy scaling argument gives a nice explanation for why the critical dimension changes. Consider the formation of a domain in an otherwise ordered phase fully breaking the maximal multipole group $\mathcal{M}^a_\max$. There are $\sim L^d$ disorder samples in the new domain. Since each sample is taken independently, the typical energy gain from forming the domain will be $\sim L^{d/2}$, by the central limit theorem. Comparing this to the cost of domain formation we calculated previously, the ordered phase will be unstable to domain nucleation when $d\le 4(a+1)$.

In clean systems, the transition from classical to quantum brought down the critical dimension because the argument depended on entropy. The Imry-Ma argument only appeals to energy considerations, so the quantum critical dimension in the presence of disorder remains the same as the classical critical dimension~\cite{Vojta2013}. In disordered quantum systems, SSB is impossible for $d\le 4(a+1)$.

@@

The Hamiltonian $H_b$ controls condensation of the site bosons. When $t_b$ is small, the site bosons are in number eigenstates while when $t_b$ is large the site bosons condense (although note that since the term proportional to $t_b$ is quartic in the site boson operators, the condensation transition may be nonstandard). Similarly, the edge bosons condense when $t_d$ is large. Note that lattice rotation symmetry will be broken if only a subset of the edge bosons condense, while if all $d$ species of condense equally then lattice rotation symmetry is preserved. We are not aware of a way to condense objects with dipolar charge while preserving {\it continuous} rotation symmetry. \ethan{when doing mean field for the dipoles the potential may be such as to favor a case where only one species condenses. The quartic part of the potential is a complicated function that depends on the dimension and parent Mott insulator, so it's better to just discuss both possibilities}

@@

\charlie{Should we say anything about Ref.~\cite{Argurio2021}, which doesn't consider multipole symmetry, but does look at restricted mobility and crystal-like phases from abnormal kinetic terms?}

In this paper we analyzed the spontaneous symmetry breaking of various multipole groups and discussed generalized Mermin-Wagner theorems, which we argued should be best understood as constraints on when a direct transition fully breaking a multipole group can occur, and when a symmetry unbroken phase and a (monopole) symmetry breaking phase must be separated by intermediate phases breaking higher multipole symmetries \ethan{is this still the best way of thinking about things? (also by ``direct transition'' I am assuming you mean ``direct continuous transition''?)}. We also considered multipole groups that are not the maximal multipole group, and the effect of quenched disorder. The disorder that we considered explicitly broke the symmetry, either fully or to a subgroup.

@@

We should be clear that the arguments in this paper are Mermin-Wagner or Imry-Ma arguments. These essentially amount to stability analyses about the Gaussian symmetry breaking fixed point. \ethan{I don't really understand how these are ``stability'' arguments --- stability with respect to which perturbations? The free theories we wrote down are at least all stable in the RG sense (the core energies of any vortex configurations can be made arbitrarily large, which means that the Gaussian actions can describe the IR physics even when the fluctuations are strong enough to prevent the symmetries from being spontaneously broken). Maybe there's something with disorder that I'm missing?}  In principle, even when the Gaussian symmetry breaking fixed point is unstable, a non-trivial fixed point with long range order could arise (see e.g. \cite{TonerRadzihovsky}). Whether and when such non-trivial fixed points can be realized in models with multipolar symmetry is an important problem for future work.

We also emphasize that when the ordered phase does not exist, we did not provide any argument for what phase should replace it. For example, in 2 spatial dimensions there can be no ordered phases of continuous monopole (ordinary) symmetries. For $O(n)$ models with $n>2$, the result is that the disordered phase is the only phase \cite{polyakov}. For $n=2$, there can in addition be quasi-long-range-ordered phases. Determining what kind of phase {\it can} obtain in the absence of long range ordered symmetry breaking is beyond the scope of this work, and would at a minimum require understanding the nature and role of topological defects in the symmetry breaking order parameter, akin to vortices in the XY model. \ethan{I think that since all the symmetries we are considering are Abelian, we can always have QLRO in the dimensions where the Gaussian action allows it (we can always work in a limit where the vortex core energy is infinitely large, meaning that we can always find a regime in which the description in terms of a phase mode is legitimate). The case of $O(n)$ models with $n>2$ is different because the sigma model for the putative broken phase is non-Abelian, with the stiffness flowing to zero. In the present Abelian case there is no such flow, and vortices are the only thing standing in the way of a Gaussian description. But as mentioned above, they can be gotten rid of (in any dimension) by making their core energies large.} The range of possible symmetry unbroken phases could be even richer in the presence of disorder, where various glassy phases could also come into play \cite{Fisheretal}. Our discussion of disorder physics was also limited to quenched short-range correlated disorder. Extensions to disorder with long-range correlations, or annealed disorder, are left to future work.

@@

{\bf Acknowledgements} This work was supported by
the U.S. Department of Energy, Office of Science, Basic
Energy Sciences, under Award \# DE-SC0021346. EL was supported by the Hertz Fellowship.

@@

\appendix

\section{No mediated condensation} \label{app:mediated}

\ethan{not sure if this appendix is necessary, given that the paper is about MW theorems?}

In Sec.~\ref{sec:partial} we saw that the generalized Mermin-Wagner argument provides critical dimensions for partial breaking that are lower than the critical dimensions for full symmetry breaking. \note{Can we use this behavior to try to evade the Mermin-Wagner argument for the full symmetry group?}

%\remove{In fact, we can use this observation to evade the generalized Mermin-Wagner theorem, completely breaking any multipole symmetry for $d>2$ classically, or $d>1$ quantum mechanically}. This possibility has been pointed out in Refs.~\cite{Griffin2015Cascade, Griffin2015}, but we include a further interpretation in terms of what types of phase transitions are allowed in different dimensions. For simplicity, we will begin by considering the classical dipole-preserving theory with symmetry group $\mathcal{M}^2_\max$.

To be explicit, consider a monopole field $\phi(x)$ and a dipole field $\varphi_i(x)$. Under the dipole symmetry they transform as 
\begin{align}
\phi(x) &\goesto \phi(x) + a_ix^i + b,\nn
\varphi_i(x) &\goesto \varphi_i(x) + a_i.
\end{align}
The leading terms in the Lagrangian will be
\begin{align}
\mathcal{L} &= (\partial_t \phi)^2 + (\partial_t \varphi_i)^2 + (\partial_i \partial_j \phi)^2 + (\partial_i \varphi_j)^2,
\end{align}
so that $\phi$ has a $k^4$ dispersion and $\varphi_i$ has a $k^2$ dispersion. We can also include an interaction term of the form $g(\partial_i \phi - \varphi_i)^2$, which does not break the symmetry.  The $\varphi_i$ field is well-defined whenever the dipole symmetry is broken.

\note{We can then redefine the $\varphi_i$ field to $\tilde{\varphi}_i = \varphi_i - \partial_i \phi$, so that it is neutral under the symmetry. The interaction term becomes $g\tilde{\varphi}_i^2$, so that the $\tilde{\varphi}_i$ field becomes massive. The only remaining massless field is $\phi$, which remains charged under the full dipole group and keeps its $k^4$ dispersion. As a result, the critical dimension for breaking the full dipole group is not affected by the possibility of partially breaking.
}
\charlie{What else to say here? Compare to Griffin/Grosvenor?}
\todo{Compare to fracton superfluids}

%As long as $\phi_i$ is well-defined, integrating out fluctuations in $\phi_i$ endows the monopole field $\phi$ with a $k^2$ dispersion. The coupling $g$ is set by how far the theory sits from the dipole critical point.

%With a $k^2$ dispersion for the monopole field, we now have a route to evade the (generalized) Mermin-Wagner argument and fully break the dipole symmetry in dimensions $d> 2$. First, partially break the symmetry so that $\phi_i$ is well defined, which is allowed for any $d>2$. Then, we can use the effective dispersion of the $\phi$ field to break the remaining part of the symmetry, also in any $d>2$. This procedure can be generalized, so that any multipole symmetry can be fully broken in $d>2$, provided we have enough symmetry-breaking order parameters. The discussion also translates to the quantum setting, so that the ultimate critical (space) dimension is always $d=1$.

%The preceding procedure gives us insight into what the generalized Mermin-Wagner arguments were telling us. When we wrote down those arguments, we did not include the intermediate fields. Thus, the generalized Mermin-Wagner arguments tell us when a direct transition between two phases is or is not allowed. Consider again the dipole group $\mathcal{M}^2_\max$. The phase in which the symmetry is fully broken can exist in any dimension $d>2$ ($d>1$ for quantum systems), but there can only be direct phase transitions between this phase and the wholly symmetry unbroken phase in $d>4$ ($d>2$). In dimensions $2<d<4$ ($1<d<2$) the fully symmetry broken and the fully symmetry unbroken phase must be separated by an intermediate phase that breaks dipole but not monopole symmetry.

>>>>

The multipole group is well-explained in Ref.~\cite{Gromov2019}. For concreteness, and since we will be interested in situations where a multipolar symmetry group is spontaneously broken, we will describe the multipole group in terms of its action on a compact scalar field $\phi$, which will play the role of a Goldstone boson for the broken multipolar symmetry. We will always imagine $\phi$ as constituting the phase mode for some microscopic order parameter $\psi \sim e^{i\phi}$, with $\psi$ transforming by various $U(1)$ phases under the multipole group. %\ethan{are there any examples showing that this interpretation is too restrictive?} RN: Not that I know of

 The multipole group generalizes the internal shift symmetry $\phi(x) \mapsto \phi (x) +c$ by allowing a shift by some set of polynomials involving the spatial coordinates, viz. $\phi (x) \mapsto \phi (x) + \lambda_\alpha P^\alpha(x)$. The variables $\lambda_\alpha$ are symmetry parameters, while $\alpha$ labels the set of polynomials $P^\alpha(x)$. These so-called polynomial shift symmetries~\cite{Griffin2015} all commute with each other. It is helpful to limit ourselves to homogeneous polynomials. We can label these as $P_a^{\alpha}$, where $a$ is the degree of the polynomial and $\alpha$ is now an index that runs over the homogeneous of degree $a$.

@@

Reference~\cite{Gromov2019} includes discussion of some multipole groups; we will review a few here. The simplest case is the maximal multipole group $\mathcal{M}^a_\text{max}$, which includes all shifts by polynomials of degree $a$ or less. Individual polynomials can be written as
\begin{align}
P_c^{\gamma} = \mu^{\gamma}_{I_c}x^{I_c}, \label{eqn:basis}
\end{align}
where $\gamma$ is an index that runs over polynomials of degree $c$ and $c\le a$. The composite index $I_c = \{i_1,\dots,i_c\}$  expands as $x^{I_c}=x^{i_1}\cdots x^{i_c}$, and similarly for $\partial_{I_c}$. Each  matrix  $\mu^{\gamma}$ is fully symmetric on its $I_c$ indices. This group also includes all translations and rotations.

@@

We can now consider building IR field theories for phases in which a multipolar symmetry is completely spontaneously broken, with no residual unbroken subgroup. Reference~\cite{Gromov2019} describes the process in detail, so in the following we will be somewhat succinct. 
We will continue to write things down in terms of the compact scalar $\phi$, which (at least in the cases for which SSB is not preempted by strong fluctuations) is to be viewed as the Goldstone for the sponatneously broken symmetry.

@@

Furthermore, the structure of the quantum correlation function,
\begin{align}
C &\sim \int \frac{d\omega d^dk}{\omega^2 + K_k}\nn
&\sim \int \frac{d^dk}{\sqrt{K_k}},
\end{align}
suggests that, at least in some broad class of theories, the quantum critical dimension will be half the classical critical dimension. This is not true in general, though, as we will show.

@@

Consider a situation in which a dipolar symmetry $\mcm^1_{\rm max}$ is spontaneously broken to its monopolar subgroup $\mcm^0$, and specialize to the case of $d=3$ and $T>0$. In the spirit of Goldstone's theorem, the most naive thing to do when analyzing the symmetry broken phase would be to write down an action involving a set of $d=3$ Goldstone fields $\theta_j$, with $\t_j$ transforming linearly under the dipolar part of the symmetry group, but remaining invariant under the unbroken monopole subgroup, so that $\theta_j \mapsto \theta_j + \beta_j$ under a transformation parameterized by $P(x) = \alpha + \beta_j x^j$. In this case, the IR field theory for the putative symmetry-breaking phase would be of the form 
\begin{align}
\label{vector} \mathcal{L}= \sum_j  (\partial_\tau \theta_j)^2 + \sum_{i,j} g_{ij} (\partial_i \theta_j)^2.
\end{align}
As we are working in $d=3$ and at $T>0$, the theory \eqref{vector} indeed has SSB for the dipole symmetry, with $e^{i\t_j}$ developing long-range order.

@@

Let $b\leq a$ denote the smallest degree of multipole charge with nonzero compressibility, i.e. consider a situation in which $\kappa_c>0$ for all $b\leq c\leq a$, while $\kappa_c=0$ for all $c<b$. Then the minimal IR Lagrangian for the symmetry broken phase looks schematically like 
\be \label{generall} \mathcal{L} = \kappa_b (\partial_\tau \phi_{I_b})^2 + g_{J_{a-b+1}}(\partial_{J_{a-b+1}}\phi_{I_b})^2,\ee 
where 
$\p_{J_{a-b+1}} \equiv \p_{j_1} \cdots \p_{j_{a-b+1}}$, and where $\phi_{I_b}$ is a field that shifts by a polynomial of degree $a-b$ under the action of $\mmax a$. The compressibility appears as the coefficient of the time derivative term since $\p_\tau \phi_{I_b}$ is the momentum conjugate to $\phi_{I_b}$ in the quantum theory, and hence correlation functions of fluctuations in $n_b$ are determined by those of $\p_\tau \phi_{I_b}$.

A necessary condition for $\mmax b$ to be spontaneously broken is that $\kappa_b \neq 0$. Indeed, if $\mmax b$ is spontaneously broken then $n_b$ creates Goldstones when acting on the ground state, thereby ensuring that the density-density correlator appearing in the calculation of \eqref{kappadef} must be nonzero at long wavelengths. % \ethan{is this too laconic?} RN Seems fine to me

@@

In the previous subsection, we saw that the number of Goldstone modes in the symmetry-breaking theory is determined by the smallest degree of multipole charge $b$ such that $\kappa_b\neq0$. We now turn to determining the symmetry breaking pattern that occurs for a given choice of $b$, spatial dimension $d,$ and temperature $T$, thereby allowing us to formulate a generalized version of the Mermin-Wagner theorem for the case of partial multipole breaking.

Since $\kappa_c=0$ if $c<b$, none of the subgroups $\mmax {c<b}$ can be spontaneously broken. Suppose then that the symmetry is spontaneously broken down to $\mmax {c}$, where $c\geq b$. To find the allowed values of $c$ consistent with this assumption, we need to calculate correlation functions of operators of the form $\exp(i\p_{J_{c-b}}\phi_{I_b})$, these being the operators transforming nontrivially under $\mmax c$ with the longest-range correlation functions. We have 
\be\ba \langle \p_{J_{c-b}}\phi_{I_b}(x) \p_{J_{c-b}}\phi_{I_b}(0)\rangle & \sim \int \frac{d\omega \, dk\, k^{d-1+2(c-b)}}{\omega^2 + k^{2(a-b+1)}} \\ 
& \sim \int dk\, k^{d-1+2(c-b) - \varsigma_T(a-b+1)}, \ea\ee 
where we have defined the function $\varsigma_T$ such that $\varsigma_T = 1$ if $T=0$ and $\varsigma_T = 2$ if $T>0$. Consequently, we find that $\mmax c$ can be spontaneously broken only in dimensions $d$ such that 
\be \label{pbreak_mw} d>2(b-c) + \varsigma_T(a-b+1).\ee 
For a given choice of symmetry breaking pattern (captured by $a$ and $c$) and a given choice of which compressibilities are nonzero (captured by $b$), the quantity on the RHS determines the lower critical dimension for symmetry breaking, thereby generalizing the Mermin-Wagner theorem to the partial breaking case. 
As a sanity check on \eqref{pbreak_mw}, note that when we set $c=b=0$ (full multipole breaking) we recover the lower critical dimension of $\varsigma_T(a+1)$ obtained in the previous section. Also note that as expected, the lower critical dimensions in the quantum $(\varsigma_T = 1)$ and classical $(\varsigma_T = 2)$ cases are related by $d_{\rm cl} = d_{\rm q} + z$, where $z=a-b+1$ is the dynamical exponent of the Goldstone theory \eqref{generall}.

@@

In a disordered classical system, the energy scaling argument gives a nice explanation for why the critical dimension changes. Consider the formation of a domain of linear size $L$ in an otherwise ordered phase fully breaking the maximal multipole group $\mathcal{M}^a_\max$. There are $\sim L^d$ disorder samples in the new domain. Since each sample is taken independently, the typical energy gain from forming the domain will be $\sim L^{d/2}$, by the central limit theorem. Comparing this to the cost of domain formation we calculated previously, the ordered phase will be unstable to domain nucleation when $d\le 4(a+1)$. We emphasize that we are here considering disorder that couples linearly to the order parameter.

In clean systems, the transition from classical to quantum brought down the critical dimension because the argument depended on entropy. The Imry-Ma argument only appeals to energy considerations, so the quantum critical dimension in the presence of disorder remains the same as the classical critical dimension~\cite{Vojta2013}. In disordered quantum systems, full SSB is impossible for $d\le 4(a+1)$.

@@

The Hamiltonian $H_b$ controls condensation of the site bosons. When $t_b$ is small, the site bosons are in number eigenstates while when $t_b$ is large the site bosons condense (although note that since the term proportional to $t_b$ is quartic in the site boson operators, the condensation transition may be nonstandard). Similarly, the edge bosons condense when $t_d$ is large. Note that lattice rotation symmetry will be broken if only a subset of the edge bosons condense, while if all $d$ species of condense equally then lattice rotation symmetry is preserved. We are not aware of a way to condense objects with dipolar charge while preserving {\it continuous} rotation symmetry. %\ethan{when doing mean field for the dipoles the potential may be such as to favor a case where only one species condenses. The quartic part of the potential is a complicated function that depends on the dimension and parent Mott insulator, so it's better to just discuss both possibilities} RN Makes sense, and ok

@@

In this paper we analyzed the spontaneous symmetry breaking of various multipole groups and discussed generalized Mermin-Wagner theorems for breaking a maximal multipole group, either fully or to a subgroup. We also considered multipole groups that are not the maximal multipole group, and the effect of quenched disorder. The disorder that we considered explicitly broke the symmetry, either fully or to a subgroup.

@@

We should be clear that the arguments in this paper are Mermin-Wagner or Imry-Ma arguments. %These essentially amount to stability analyses about the Gaussian symmetry breaking fixed point. %\ethan{I don't really understand how these are ``stability'' arguments --- stability with respect to which perturbations? The free theories we wrote down are at least all stable in the RG sense (the core energies of any vortex configurations can be made arbitrarily large, which means that the Gaussian actions can describe the IR physics even when the fluctuations are strong enough to prevent the symmetries from being spontaneously broken). Maybe there's something with disorder that I'm missing?} RN: Ethan, if stuff is stable, there is no problem. What this discussion was intended to say, is that even if the Gaussian symmetry breaking fixed point is unstable to fluctuations, you could have a non-trivial but still symmetry breaking fixed point, and that is beyond the analysis. is the new discussion clearer? 
In principle, even when the Gaussian symmetry breaking fixed point is unstable to fluctuations, a non-trivial fixed point with long range order could arise (see e.g. \cite{TonerRadzihovsky}). Whether and when such non-trivial fixed points can be realized in models with multipolar symmetry is an important problem for future work.

We also emphasize that when the ordered phase does not exist, we did not provide any argument for what phase should replace it. For example, in 2 spatial dimensions there can be no ordered phases of continuous monopole (ordinary) symmetries. For $O(n)$ models with $n>2$, the result is that the disordered phase is the only phase \cite{polyakov}. For $n=2$, there can in addition be quasi-long-range-ordered phases. Determining what kind of phase {\it can} be obtained in the absence of long range ordered symmetry breaking would at a minimum require understanding the nature and role of topological defects in the symmetry breaking order parameter, akin to vortices in the XY model. We expect that if we stick with an $O(2)$ internal group, then quasi-long range order should always be a possibility at the marginal spatial dimension, at least in the limit where the vortex core energy is infinitely large. The extension to non-Abelian internal groups is beyond the scope of this work. %\ethan{I think that since all the symmetries we are considering are Abelian, we can always have QLRO in the dimensions where the Gaussian action allows it (we can always work in a limit where the vortex core energy is infinitely large, meaning that we can always find a regime in which the description in terms of a phase mode is legitimate). The case of $O(n)$ models with $n>2$ is different because the sigma model for the putative broken phase is non-Abelian, with the stiffness flowing to zero. In the present Abelian case there is no such flow, and vortices are the only thing standing in the way of a Gaussian description. But as mentioned above, they can be gotten rid of (in any dimension) by making their core energies large.} RN: Accepted Ethan's comment and incorporated it into the text
The range of possible symmetry unbroken phases could also be even richer in the presence of disorder, where various glassy phases could also come into play \cite{Fisheretal}. Our discussion of disorder physics was also limited to quenched short-range correlated disorder. Extensions to disorder with long-range correlations, or annealed disorder, are left to future work.

@@

{\bf Acknowledgements} This work was supported by
the U.S. Department of Energy, Office of Science, Basic
Energy Sciences, under Award \# DE-SC0021346. E.L. was supported by a Hertz Fellowship.

@@
